---
title: "Class 5: Data Visualization with ggplot"
author: "Andrew Quach (A12137777)"
format: pdf
editor: visual
---

##Using GGPLOT

To use ggplot2 we first need to install it on our computers. To do this we will use the function `install.packages()`

Before I use any package functions I have to load them up with a `library()` call, like so:
```{r}
library(ggplot2)
head(cars)
```
```{r}
cars
```
There is always the "base R" graphics system, i.e. `plot()`

```{r}
plot(cars)
```
To use ggplot I need to spell out at least 3 sthing: 
-data(the stuff I want to plot as a data.frame)
-aes (aes() values - how the data map to the plot)
-geoms ( how I want things drawn)

```{r}
ggplot(cars) +
  aes(x=speed, y=dist)
```
```{r}
ggplot(cars) +
  aes(x=speed, y=dist) +
  geom_point()
```


```{r}
ggplot(cars) +
  aes(x=speed, y=dist) + 
  geom_point() +
  geom_smooth(method="lm")
```

```{r}
ggplot(cars) +
  aes(x=speed, y=dist) + 
  geom_point() +
  geom_smooth(method="lm", se=FALSE)
```

```{r}
ggplot(cars) + 
  aes(x=speed, y=dist) +
  geom_point() +
  labs(title="Speed and Stopping Distances of Cars",
       x="Speed (MPH)", 
       y="Stopping Distance (ft)",
       subtitle = "Your informative subtitle text here",
       caption="Dataset: 'cars'") +
  geom_smooth(method="lm", se=FALSE) +
  theme_bw()
```

```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```
```{r}
nrow(genes)
```
```{r}
colnames(genes)
```
```{r}
ncol(genes)
```
```{r}
table(genes$State)
```
```{r}
round ( table(genes$State)/nrow(genes) * 100, 2)
```
```{r}
ggplot(genes) + 
  aes(Condition1, Condition2) + 
  geom_point()
```
```{r}
p <- ggplot(genes) +
  aes(x=Condition1, y=Condition2, col=State) +
  geom_point()
p
```

```{r}
p + scale_colour_manual( values=c("blue", "gray", "red") )
```

```{r}
p + scale_colour_manual(values=c("blue", "gray", "red")) +
  labs(title="Gene Expression Changes Upon Drug Development", x="Control (no drug)", y="Drug Treatment")
```


```{r}
url <- "https://raw.githubusercontent.com/jennybc/gapminder/master/inst/extdata/gapminder.tsv"

gapminder <- read.delim(url)
```

```{r}
# install.packages("dplyr")  ## un-comment to install if needed
library(dplyr)

gapminder_2007 <- gapminder %>% filter(year==2007)
```

```{r}
ggplot(gapminder_2007) +
  aes(x=gdpPercap, y=lifeExp) + 
  geom_point()
```

```{r}
ggplot(gapminder_2007) + 
  aes(x=gdpPercap, y=lifeExp) +
  geom_point(alpha=0.5)
```

```{r}
ggplot(gapminder_2007) +
  aes(x = gdpPercap, y = lifeExp, color = pop) +
  geom_point(alpha=0.8)
```

```{r}
ggplot(gapminder_2007) + 
  aes(x = gdpPercap, y = lifeExp, size = pop) +
  geom_point(alpha=0.5)
```

```{r}
ggplot(gapminder_2007) + 
  geom_point(aes(x = gdpPercap, y = lifeExp,
                 size = pop), alpha=0.5) + 
  scale_size_area(max_size = 10)
```

```{r}
gapminder_1957 <- gapminder %>% filter(year==1957)

ggplot(gapminder_1957) + 
  aes(x = gdpPercap, y = lifeExp, color=continent,
                 size = pop) +
  geom_point(alpha=0.7) + 
  scale_size_area(max_size = 10) 
```

```{r}
gapminder_1957 <- gapminder %>% filter(year==1957 | year==2007)

ggplot(gapminder_1957) + 
  geom_point(aes(x = gdpPercap, y = lifeExp, color=continent,
                 size = pop), alpha=0.7) + 
  scale_size_area(max_size = 10) +
  facet_wrap(~year)
```

